---
title: "ms_3"
author: "Lindsey Greenhill"
date: "02/28/2020"
output: html_document
---
## About
The purpose of this project is to analyze a data set compiled by the Counter
Trafficking Data Collaborative. This data set is an anonymized collection of
reports of human trafficking victims across the globe from 2002 to 2017. The
data set contains information on about 47,000 victims. The observations in this
data set primarily come from a standardized combination of Polaris and IOM data.
This data is not necessarily intended to be a representative sample of all human
trafficking victims, but is one of the only large and reliable data sets on
human trafficking publicly available. I chose to only look at data past 2010,
because much of the data before then is missing. I plan on primarily focusing on
the country of origin and country of destination for each trafficking victims. I
am currently planning on using the polity index (which measures the level of
democracy in a country and periods of transition) in conjunction with the ctdc
data to see if there are any interesting patterns. I am also planning on
focusing on non-US human trafficking, because American trafficking operates very
differently than European, African, and Asian trafficking. I have joined the
ctdc data and polity data based on country name and year.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(janitor)
library(gt)
library(tidyverse)
```
```{r ctdc_data, include=FALSE}

# reading in data from the Counter Trafficking Data Collective with 62 different
# variables. Data set made in 2017. Includes 47,000 different victims of 43
# different nationalities, with exploitation taking place across 52 different
# countries. Original data class included 80,000 observations, but after
# k-anonymization widdled down. Data set compiled through analysis of different
# data models from different contributors. Polaris and IOM cofounders of this
# project. Data set not random, not necessarily representative sample of all
# human trafficking victims. Data from Liberty Asia also included. Years frange
# from 2002 to 2017. want to keep yearOfRegistration, datasource,
# gender,majorityStatus, majorityStatusAtExploit (-99 is NA), citizenship (need
# to link with values based on iso country codes), all variables starting with
# "is", include all "typeOfLabour" variables. Might not do all of these
# initially.

ctdc_data <- read_csv("raw-data/the_global_dataset_3_sept_2018.csv", na = c("", "-99", "NA")) %>% 
  clean_names() %>%
  select(year_of_registration, datasource, gender,
         majority_status, citizenship, country_of_exploitation,
         ) %>%
    filter(year_of_registration >= 2010, citizenship != "US")
```

```{r country_code_data, include=FALSE}

# data from iso on country codes, which are used in teh ctdc data set

code_data <- read_csv("raw-data/data_csv.csv")  %>% clean_names()
```

```{r polity_data, include=FALSE}

#reading in polity data. filtering for rows where year is >= 2010, because I'm
#only using data from then. Code book for this data set important to use.
#Country codes three characters instead of two, so I will try to join based on
#the country name instead. If that doesn't work I'll find a set that converts
#from iso to alpha codes. Want to keep: country, year, polity, durable, persist

polity_data <- read_excel("raw-data/polity.xls") %>%
  clean_names() %>%
  filter(year >= 2010) %>%
  select(country, year, polity, durable)
polity_data

```

```{r join_data, include=FALSE}

# joined by iso country codes. Had to use inner_joins because it wasn't working
# in just one. I also tried to use the suffix argument, but that  wouldn't work
# because I would have needed to inner join by the two variables at the same
# time, so I used mutate instead.

joined_country_code_data <- ctdc_data %>%
  inner_join(code_data, by = c("citizenship" = "code")) %>%
  inner_join(code_data, by = c("country_of_exploitation" = "code"))  %>%
  rename(origin_country = name.x, destination_country = name.y) %>%
  select(-majority_status, - citizenship, -country_of_exploitation, -datasource)
joined_country_code_data
```

```{r join_polity_data, include = FALSE}

# joining based on year and country name. Had the same problem joining by two
# variables as above so I had to use two inner joins. renamed the columns also
# because suffix argument not working.

joined_ctdc_code_polity <- joined_country_code_data %>%
    inner_join(polity_data, by = c("year_of_registration" = "year", "origin_country" = "country")) %>%
  inner_join(polity_data, by = c("year_of_registration" = "year", "destination_country" = "country")) %>%
  rename(polity_origin = polity.x, durable_origin = durable.x, polity_destination = polity.y, durable_destination = durable.y, year = year_of_registration)

```
```{r show_table, echo=FALSE}
joined_ctdc_code_polity %>% slice(1:20) %>%
  gt() %>%
  tab_header(title = "CTDC and Polity Joined Data")
```
Repo url: https://github.com/lindseygreenhill/final-project 

