---
title: "ms_4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(janitor)
library(gt)
library(tidyverse)
```

## Thoughts

I feel like I pretty thoroughly explained my data in the last milestone, so I
thought I would discuss in more depth the relationships I hope to investigate
using this data. My vision for this project is to do an exploration of the CTDC
data. As of now, I have looked at the most common origin and destination
countries. I am thinking of creating a dodged bar chart of the frequencies of
these countries which would give a good comparison I think. I am then going to
look at the top five origin countries and looking at the characteristics of
their destination countries. For example, what polity level do their destination
countries have? what gdp? I have included some preliminary graphs that I have
made for cambodia and the Philippines investigating their destination countries.
I have yet to find any strong relationship with the polity scores, but I will
continue looking. I'm also currently trying to find easy GDP data for the
countries by year and not just in 2020 or 2019. I need it dating back to 2010.

```{r ctdc_data, include=FALSE}

# reading in data from the Counter Trafficking Data Collective with 62 different
# variables. Data set made in 2017. Includes 47,000 different victims of 43
# different nationalities, with exploitation taking place across 52 different
# countries. Original data class included 80,000 observations, but after
# k-anonymization widdled down. Data set compiled through analysis of different
# data models from different contributors. Polaris and IOM cofounders of this
# project. Data set not random, not necessarily representative sample of all
# human trafficking victims. Data from Liberty Asia also included. Years frange
# from 2002 to 2017. want to keep yearOfRegistration, datasource,
# gender,majorityStatus, majorityStatusAtExploit (-99 is NA), citizenship (need
# to link with values based on iso country codes), all variables starting with
# "is", include all "typeOfLabour" variables. Might not do all of these
# initially.

ctdc_data <- read_csv("raw-data/the_global_dataset_3_sept_2018.csv", na = c("", "-99", "NA")) %>% 
  clean_names() %>%
  select(year_of_registration, datasource, gender,
         majority_status, citizenship, country_of_exploitation,
         ) %>%
    filter(year_of_registration >= 2010, citizenship != "US")
```

```{r country_code_data, include=FALSE}

# data from iso on country codes, which are used in teh ctdc data set

code_data <- read_csv("raw-data/data_csv.csv")  %>% clean_names()
```

```{r polity_data, include=FALSE}

#reading in polity data. filtering for rows where year is >= 2010, because I'm
#only using data from then. Code book for this data set important to use.
#Country codes three characters instead of two, so I will try to join based on
#the country name instead. If that doesn't work I'll find a set that converts
#from iso to alpha codes. Want to keep: country, year, polity, durable, persist

polity_data <- read_excel("raw-data/polity.xls") %>%
  clean_names() %>%
  filter(year >= 2010) %>%
  select(country, year, polity, durable)
polity_data

```

```{r join_data, include=FALSE}

# joined by iso country codes. Had to use inner_joins because it wasn't working
# in just one. I also tried to use the suffix argument, but that  wouldn't work
# because I would have needed to inner join by the two variables at the same
# time, so I used mutate instead.

joined_country_code_data <- ctdc_data %>%
  inner_join(code_data, by = c("citizenship" = "code")) %>%
  inner_join(code_data, by = c("country_of_exploitation" = "code"))  %>%
  rename(origin_country = name.x, destination_country = name.y) %>%
  select(-majority_status, - citizenship, -country_of_exploitation, -datasource)
joined_country_code_data
```

```{r join_polity_data, include = FALSE}

# joining based on year and country name. Had the same problem joining by two
# variables as above so I had to use two inner joins. renamed the columns also
# because suffix argument not working.

joined_ctdc_code_polity <- joined_country_code_data %>%
    inner_join(polity_data, by = c("year_of_registration" = "year", "origin_country" = "country")) %>%
  inner_join(polity_data, by = c("year_of_registration" = "year", "destination_country" = "country")) %>%
  rename(polity_origin = polity.x, durable_origin = durable.x, polity_destination = polity.y, durable_destination = durable.y, year = year_of_registration)

```
## Cambodia Graph 
```{r Cambodia, echo=FALSE}

# data set with just Cambodia information. Plot showing destination countries

Cambodia <- joined_ctdc_code_polity %>%
  filter(origin_country == "Cambodia") 
Cambodia_destinations <- ggplot(Cambodia, aes(destination_country)) +
  geom_bar() +
  labs(title = "Cambodia's Destination Countries")
Cambodia_destinations

```

## Philippines Graph

```{r Philipinnes, echo=FALSE}

# data set identical to Cambodia but with Philippines. Plot showing destination
# countries

Philippines <- joined_ctdc_code_polity %>%
  filter(origin_country == "Philippines") 
Philippines_destinations <- ggplot(Philippines, aes(destination_country)) +
  geom_bar() +
  labs(title = "Philippines's Destination Countries")
Philippines_destinations
```
